---
title: "ASSESSMENT 1"
author: "Shuvomita Biswas (S4669000)"
date: "16/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#RMarkdown editing
R Markdown documents provide  ** quick, reproducible ** reporting from R. You write your document in markdown and embed executable R code chunks with the knitr syntax. You can update your document at any time by *re-knitting the code chunks*. You can then convert your document into several common formats.

5. Review the suggested documentation on how to insert links in the .rmd file and include an in-text link to https://rmarkdown.rstudio.com/lesson-8.html 

6. Insert an R chunk and create a variable with this dataset
https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv

The name of the variable should be `fastfood`

---------------------------------------------------------------------------------------------------------------------
```{r}
library(tidyverse)
library(knitr)
library(skimr)
```
```{r}
fastfood <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv")
```
---------------------------------------------------------------------------------------------------------------------
```
7. Display the first 10 rows of the dataset using head() and kable(). 

Display the first 10 rows of the dataset and the first 5 variables

Use Help and the link below to read more about those functions  https://bookdown.org/yihui/rmarkdown-cookbook/kable.html

```{r}
kable (fastfood [1:10,], caption = "First 10 rows")
head(fastfood, format = "First 10 rows")
```
8. Display the observations that has more than 1000 calories

```{r}
fastfood %>%
  filter(calories > 1000)
```
9. Arrange observations with more than 40 in total_fat and more than 80 in total_carb in the descending order and save them to a new variable (dataset) called `dont_eat_this'

Total_fat and total_card cannot be arrange in descingding order at the same time
```{r}
dont_eat_this<-fastfood
dont_eat_this %>%
  filter(total_fat>40,total_carb>80)%>%
  arrange(desc(total_fat,total_carb))
```

10. Using the initial dataset variable, use `case_when()` to create a new variable `heavy_food` which is equal to "heavy" when total calories are greater than 500, "low" when total calories are less than 250 and "average" for all other cases. Count the number of observations in each created category.
```
```{r}
heavy_food <- fastfood
heavy_food %>%
  mutate(heavy_food = case_when (calories > 500 ~ "heavy",
                                 calories < 250 ~ "low",
                                 T ~ "average"))
```
11.Display the types of variables in the dataset using `skimr` package 

```{r}
skim(heavy_food)
```
12.Present the count observations from each restaurant in a descending order

Show the number of distnct items on a menu in the dataset

```{r}
fastfood%>%
  count(restaurant)%>%
  arrange(desc(n))
fastfood %>%
  n_distinct(fastfood $ item)
```

13. Using groupings (group_by()), summarise and display the average number of calories for each restaurant.

```{r}
fastfood %>%
  group_by(calories)%>%
  summarise(average_calories = mean(calories)) %>%
  ungroup()
```

14. Add variables to the dataset, which:

-  calculates the average calories per type of restaurant and call it `average_calories` 
-  calculates the maximum total_fat per type of restaurant and call it `max_fat` 
-  calculates the minimum cholesterol per type of restaurant and call it `min_cholesterol` 

```{r}
fastfood %>%
  group_by(restaurant)%>%
  summarise(average_calories = mean(calories)) %>%
  ungroup()

fastfood %>%
  group_by(restaurant)%>%
  summarise(max_fat = max(total_fat)) %>%
  ungroup()

fastfood %>%
  group_by(restaurant)%>%
  summarise(min_cholesterol = min(cholesterol)) %>%
  ungroup()
```

15.Display the data vis of total fat per each type of restaurant. Write a narration (2-3 sentences) why you believe this type of data viz presents such information best. 

This visualisation is used for portraying functions in which restaurant is the independent variable.As I have used point graph, it would be easier for the reader to view the data clearly.
```{r}
ggplot(fastfood, aes(restaurant, total_fat)) +
  geom_point() +
  coord_flip()
```
16. Add a variable to the dataset, which calculates the sum of cholesterol and sodium and call it `cholesterol_sodium`.

Remove the variable `salad`

```{r}
cholesterol_sodium <- fastfood
cholesterol_sodium %>%
  mutate(cholesterol_sodium = sum(cholesterol,sodium)) %>%
  select(-salad)
```
17. Use observations for Mcdonalds to plot sugar variable against protein with `geom_point()` 

```{r}
ggplot(filter(fastfood, restaurant == "Mcdonalds"), aes(x=sugar, y=protein, colour())) + geom_point()
```

18.Identify variable(s) which should be factors and transform their type into a factor variable.


```{r}
str(fastfood)
fastfood $ restaurant <- as.factor(fastfood $ restaurant)
skim(fastfood)
```

19. Create two new variables:

Read about `cut_number()` function using Help and add a new variable to the dataset `calories_type`. Use `calories` variable for `cut_number()` function to split it into 3 categories `n=3`, add labels `labels=c("low", "med", "high")` and make the dataset ordered by arranging it according to calories. 


```{r}
calories_type <- fastfood 
calories_type %>%
  mutate(calories_type = cut_number(calories,n=3,labels = c("low","med","high")))%>%
  arrange(calories)
```

20. Create a dataviz that shows the distribution of `calories_type` in food items for each type of restaurant. Think carefully about the choice of data viz. Use facets, coordinates and theme layers to make your data viz visually appealing and meaningful. Use factors related data viz functions.

```{r}
calories_type %>%
count(calories, restaurant) %>%
ggplot(aes(restaurant,calories)) +
geom_point(colour = "blue") +
coord_flip()
 
```

21.The second variable should show the percentage of `trans_fat` in `total_fat`. Add the variable to the dataset and call it `trans_fat_percent`. Do not forget to save the updated dataset. 

```{r}
trans_fat_percent <- fastfood  
trans_fat_percent %>%
  mutate(trans_fat_percent = trans_fat/total_fat * 10)
```

22. Create a dataviz that shows the distribution of `trans_fat` in food items for each type of restaurant. Think carefully about the choice of data viz. Use facets, coordinates and theme layers to make your data viz visually appealing and meaningful.


```{r}
ggplot(fastfood, aes(restaurant, trans_fat, fill = restaurant)) + 
  geom_col() +
  theme_linedraw()
```

23.Calculate and show the average (mean)  `total_fat` for each type of restaurant. No need to save it as a variable.

```{r}
fastfood %>% group_by(restaurant) %>%
  summarise(total_fat = mean(total_fat))
```

24. And create a dataviz that allow to compare different restaurants on this variable (`total_fat`). You can present it on one dataviz (= no facets). 

```{r}
ggplot(fastfood, aes(restaurant, total_fat)) +
  geom_point(colour= "red") +
  theme_linedraw()

```

